Supplemental Material for ``Assessing United States county-level exposure for research on tropical cyclones and human health"

### 2020-08-17 ehp_revision/response_to_reviewers.Rmd

\begin{shaded}
\textbf{R2 Comment 5:}
Table 2: This is another confusing table. Not sure what compared here. The
second column is a mean \# of exposed counties over years and the third column
seems like showing single storm with the max \# of exposed counties. Not sure
what info I should digest from this table.
\end{shaded}

> **Text from Results of original manuscript:**
"Across the four storm hazards considered, there was wide variation in the
average number of county exposures per year (Table 2). For tropical cyclone
tornadoes, there were on average about 40 county exposures per year within our
study. County exposures were more frequent for tropical cyclone wind exposures
(>160/year on average), even more frequent for tropical cyclone flood exposures
(>190/year on average), and most frequent for tropical cyclone rain exposure
(>290/year on average). For every hazard except tornadoes, we identified at
least one tropical cyclone that exposed over 250 counties (Table 2). However,
the largest-extent tropical cyclone varied across hazards: Frances in 2004
exposed the most counties based on rain, Michael in 2018 based on wind, and Ivan
in 2004 based on flooding and tornadoes (Table 2)."

**Response:**

We hope that these results can help epidemiologists who are planning new
multi-year studies of tropical cyclones and associated health risks. For 
planning a multi-year study, it is helpful to know how many exposures the 
study might incorporate, as this can help in giving a general idea of whether
the study might have adequate statistical power to investigate a certain 
research question. This table provides some basic summaries of the extent
of exposure to different tropical cyclone hazards, both on average for the
storms in the study and also for those storms that were most extensive. 
Further, as with other results in the study, it provides evidence that 
patterns in different hazards can vary substantially, both because some 
hazards are much less likely to expose counties (e.g., torndaoes are a fairly 
rare exposure, based on this table) and also because there is variation in 
the most extensive storms across the hazards considered.

We have added text to help the readers interpret this table ...

### 2020-08-17 ehp_revision/response_to_reviewers.Rmd

> **Text in revised manuscript:**
"While other approaches have been developed to incorporate storm
hazards, particularly wind, into exposure assessment (e.g., Grabich et al.
2015a; Zandbergen 2009; Czajkowski et al. 2011), there is not yet a standard
approach. When different studies use different datasets or storm hazards
when assessing storm exposure, it becomes difficult to compare and aggregate
findings."

\begin{shaded}
\textbf{R1 Comment 6:}
This is not a comment on the paper, but the authors may also want to consider
publishing their work in DesignSafe, which is "the web-based research platform
of the NHERI  [NSF-Funded Natural Hazards Engineering Research Infrastructure]
Network that provides the computational tools needed to manage, analyze, and
understand critical data for natural hazards research." It is the data
repository widely used by interdisciplinary hazards and disaster researchers.
\end{shaded}

**Response:**

It turns out that, because we have shared the data as an R package, it should
already be accessible to use on the DesignSafe platform.

DesignSafe is an online platform sponsered by the National Science Foundation as
part of the Natural Hazards Engineering Research Infrastructure
[@rathje2017designsafe]. This platform can be used to store, share, and publish
citable data related to natural hazards through its Data Depot. It goes beyond a
data repository, however, and also allows researchers to conduct data
integration and analysis in the cloud through its Discovery Workspace, while
drawing on data stored in the Data Depot and while using scientific programming
tools.

As part of its platform, DesignSafe allows users to work in a number of
programming languages, including R (through Jupyter Notebooks)
[@rathje2017designsafe]. Therefore, DesignSafe users already have access to any
data shared through an R package, including the data we present in this
manuscript, without the data needing to be separately loaded and published
through DesignSafe's Data Depot.

This is great, because it might otherwise be difficult to try to maintain copies
of the dataset in different repositories (i.e., one copy as an R package and
another as a DesignSafe dataset), and there could be confusion about things like
the equivalence of the two copies. R packages are citable through a standard
format that gives the R package version, while data publicly shared on
DesignSafe is assigned a digital object identifiers (DOI)
[@rathje2017designsafe], and so different citations would be used for the same
dataset depending on how the user accessed it. If updates of one copy were not
immediately updated for the other copy, the copies could get out of sync, and
maintaining two synced copies would add would add an extra layer of work with
each update to the R package (we have been updating several hazards as new data
becomes available). DesignSafe's interface with R avoids this problem by
allowing its users direct access to our package and its data through that path,
as it allows for a single, citable, licensed version of the data stored and
shared from a single location, with no question of syncing separate versions of
the data or clarifying licensing and citation practices across two storage
locations.

We have added a note in the manuscript to let readers know that, because the
data is shared through an R package, and because DesignSafe's Discovery
Workspace allows users to work in R, the data can be fully accessed and used for
those working on the DesignSafe platform by using the same code in a DesignSafe
Jupyter notebook that would otherwise be used on a local computer.

> **Text added to the Methods of the revised manuscript (added text in bold):**
"**By sharing this data as an R package, it is also accessible through 
cloud-based computing platforms that incorporate Jupyter notebooks, including
the National Science Foundations's DesignSafe platform for natural hazards
engineering research [@rathje2017designsafe].**"

@@

\textit{Lines 185--187 from original manuscript:}

\begin{quotation}
\noindent
\textit{"We used tracking data from HURDAT2, which records the storm center’s
position at the synoptic times of 6:00 am, 12:00 pm, 6:00 pm, and 12:00 am
Coordinated Universal Time (UTC)."}
\end{quotation}
\end{shaded}

**Response:**

Some weather data are collected at standardized times, so that observations from
different locations can be meaningfully integrated to provide a view of
large-scale (synoptic) weather systems and patterns [@usnwstime;
@willoughby2007hurricane; @willis2006cleveland; @raines1996getting]. This
practice extends back over 100 years, when weather data for the US were
collected by local weather offices and sent by telegram at regular times each
day into a central office, which used this information to create synoptic
weather maps that captured major weather systems in the county and from this
sent back forecasts to local weather offices, also at standardized times
[@willis2006cleveland; @raines1996getting]. Synoptic weather times are typically
set based on the Universal Time Constant, with times recorded in Coordinated
Universal Time (i.e., Greenwich Mean Time), and time stamps on weather maps and
data based on this time zone often is given a "Z" (for "Zulu", an earlier name
for this timezone) after the time [@usnwstime;
@morris2008time].

@@

\begin{quotation}
\noindent
\textit{"To identify flood- and tornado-based tropical cyclone exposures in US
counties, we matched storm tracks with event listings from the National Oceanic
and Atmospheric Administration (NOAA)’s Storm Event Database (NOAA NCEI 2020).
While this database has recorded storm data, particularly tornadoes, since 1950,
its coverage changed substantially in 1996 to cover more types of storm events,
including flood events (NOAA NCEI 2020). We therefore only considered flood
metrics of tropical cyclone exposure for storms in 1996 and later."}
\end{quotation}
\end{shaded}

**Response:**

As suggested, we have added that these events are based on reports in the
Methods, in addition to our previous discussion of the point in the Discussion:

> **Revised text (relevant addition in bold):**
To identify flood- and tornado-based tropical cyclone exposures in US counties,
we matched storm tracks with event listings from the National Oceanic and
Atmospheric Administration (NOAA)’s Storm Event Database (NOAA NCEI 2020).
**Events are listed in this database based on reports from National Weather
Service personnel and other sources.** While this database has recorded storm
data, particularly tornadoes, since 1950, its coverage changed substantially in
1996 to cover more types of storm events, including flood events (NOAA NCEI
2020). We therefore only considered flood metrics of tropical cyclone exposure
for storms in 1996 and later."

### 2020-08-17 ehp_revision/writing/ehp_methods.tex
Work more on response

To identify flood- and tornado-based tropical cyclone exposures in \ac{US}
counties, we matched storm tracks with event listings from the National Oceanic
and Atmospheric Administration (NOAA)'s Storm Event
Database~\parencite{stormevents}. While this database has recorded storm data,
particularly tornadoes, since~1950, its coverage changed substantially in~1996
to cover more types of storm events, including flood
events~\parencite{stormevents}. We therefore only considered flood metrics of
tropical cyclone exposure for storms in~1996 and later.

@@

We have posted most study data as an open-source R package
\parencite{hurricaneexposuredata}. We have posted remaining data and code at
\url{https://github.com/geanders/county_hurricane_exposure}.

>>>>

To identify flood- and tornado-based tropical cyclone exposures in \ac{US}
counties, we matched storm tracks with event listings from the National Oceanic
and Atmospheric Administration (NOAA)'s Storm Event
Database~\parencite{stormevents}. \textbf{Events are listed in this database
based on reports from National Weather Service personnel and other sources.}
While this database has recorded storm data, particularly tornadoes,
since~1950, its coverage changed substantially in~1996 to cover more types of
storm events, including flood events~\parencite{stormevents}. We therefore only
considered flood metrics of tropical cyclone exposure for storms in~1996 and
later.

@@

We have posted most study data as an open-source R package
\parencite{hurricaneexposuredata}. \textbf{By sharing this data as an R package, it
is also accessible through cloud-based computing platforms that
incorporate Jupyter notebooks, including the National Science Foundations's
DesignSafe platform for natural hazards engineering research
\parencite{rathje2017designsafe}.} We have posted remaining data and code at
\url{https://github.com/geanders/county_hurricane_exposure}.

### 2020-08-17 ehp_revision/writing/ehp_response_darlings.tex
Work more on response

>>>>

# R1 Comment 6

*What is DesignSafe:*

> "DesignSafe is the cyberinfrastructure platform that has been developed as
part of NHERI to support natural hazards engineering research, and
it succeeds the NEEShub cyberinfrastructure that was developed for the
earthquake engineering community through the Network for Earthquake
Engineering Simulation (NEES) program." [@rathje2017designsafe]

> "DesignSafe plays an important role in integrating the various NHERI
components and the research taking place at NHERI facilities, but also has the
broader goal of enabling transformative research in natural hazards and
engineering across the numerous technical disciplines engaged in this field."
[@rathje2017designsafe]

> "DesignSafe has been developed as a flexible, extensible, community-driven
cyberinfrastructure, and it embraces a cloud strategy for the big data generated
in natural hazards engineering. DesignSafe provides a comprehensive CI that
supports the full research lifecycle from planning to execution to analysis to
publication and curation." [@rathje2017designsafe]

> "The future of natural hazards engineering research requires the integration
of diverse data sets from a variety of sources including experiments,
computational simulation, and field reconnaissance, as well as a variety of
research disciplines including earth science, social science, building science,
and architecture. The DesignSafe cyberinfrastructure has been designed to
provide the functionalities that will enable transformative research in natural
hazards engineering. By adopting a cloud strategy, DesignSafe allows for a
fundamental change in the way that research is performed. It provides a
comprehensive cyberinfrastructure that supports research workflows, data
analysis, and visualization, as well as the full lifecycle of experimental,
field, and computational research required by engineers and scientists to
effectively address the threats posed to civil infrastructure by natural
hazards. The integration of data and computation in the cloud will enable new
research discoveries in natural hazards engineering, which in turn can lead to
more hazard-resilient civil infrastructure." [@rathje2017designsafe]

*Access to R:*

> "The real revolution takes place downstream of data generation. Here, data 
analytics and visualization are performed in the cloud within the *Discovery
Workspace* while accessing any data within the *Data Depot*. Researchers can 
invoke common analysis programs, such as MATLAB, as well as other analysis/
visualization tools, such as Jupyter notebooks. A Jupyter notebook is an 
electronic notebook that allows users to embed rich text elements as well
as computer code, graphs, and visualizations within a single notebook that
can be shared through the web. Over 40 different programming languages are 
supported in Jupyter, including Python and R, and MATLAB code can be easily 
converted, making Jupyter a versatile tool for research. Performing analysis in 
the cloud allows researchers to integrate and explore various data without
tedious downloads. Additionally, using a seamless cyberinfrastructure to 
complete all research tasks enables tracking and relating of the processes 
applied to data." [@rathje2017designsafe]

*Data Depot:*

> "the Data Depot, a flexible data repository with streamlined data management
tools." [@rathje2017designsafe]

> "As data progresses towards publication the requirements for metadata 
increase as metadata provides users with search and discovery functions. At
the end of the research project the user may edit the information for 
publication and complete the process of assigning digital object identifiers
(DOIs) and applying the appropriate license. On-demand assistance from a 
curator is available to provide training and guide users through their data
curation and publication needs." [@rathje2017designsafe]

# R3 Comment 2

"Synoptic meteorology" refers to ...
A key idea is that observations from a widespread geographical area can be
integrate to provide a picture of large-scale weather [features?], including
large areas of high or low pressure [?].

> "In meteorology, the term synoptic denotes simultaneous observations taken
globally, or at least over a large area." [@willoughby2007hurricane]

> "Nearly 20 years ago [from 1937] [Bjerknes] declared that numerous and 
important problems might be solved with the aid of observations already 
assembled in the archives of different bureaus, but that the difficulties
in making these observations serve a special purpose were unsurmountable
for the individual scholar. He added that, in particular, the synoptic 
charts representing the momentary states of the atmosphere over large
areas of the earth, in all the necessary details and exactitude, were 
excessively difficult to assemble. His proposal was that there be entered
at the central bureau of each country, on the appropriate maps, all the available
observations taken in that country and that there be employed, by the different
bureaus, maps that may be superimposed or juxtaposed." [@gregg1937international]

However, to integrate data from many sources to provide inputs for broadscale,
synoptic analysis, it is helpful to use some standards in collecting these data
at different locations. To facilitate numerical weather forecasting and other
synoptic meteorology, weather data are regularly collected at standardized
times, so that the data can be more easily integrated as an input to models.
This practice extends back over 100 years, when data were collected from many
[offices?] in the US by telegram [?] at regular times during the day and used to
create synoptic weather maps that captured major weather systems in the country.

> "In order to allow comparison of surface conditions at the same time
throughout the world, the groups regulating the collection of
hydrometeorological data have established a standard time for taking
observations. These observations will have times observed within the 10 minutes
preceding the UTC standard time. Hence surface observations for meteorological
use have timestamps reflecting 0000, 0300, 0600, 0900, 1200, 1500, 1800, 2100
UTC." [@usnwstime]

> "Since the end of World War II, observers have launched rawinsonde observation
worldwide at 00 and 12 UTC (coordinated universal time). High-threat situations,
such as impending TC landfall, may dictate observations at 6-h or even 3-h
intervals. Interpolating rawinsonde observations to the model's mesh and
starting the calculation at a synoptic time is the simplest way to prepare a
model initial condition." [@willoughby2007hurricane]

> "Because simple interpolation in space and time is a less-than-optimum way to
use observations, nonsynoptic data require elaborate four-dimensional
variational data assimilation schemes." [@willoughby2007hurricane]

> "By 1941, charts for the 750, 500, and 250 mbar surfaces were being prepared
daily for 0400, 1200, and 2000 GMT." [@grahame2000development]

> "During the period considered here, NHC issued advisories 3 h after the
standard synoptic time (0300, 0900, 1500, and 2100 UTC)." [@torn2012uncertainty]

> "The operation data assimilation and forecast cycle interval is 6 h"
[@aberson201010]

> "Dan Petersen asked whether a data deficiency really exists with all the
observational sources in the Atlantic region, or are the data not being properly
used in the numerical models? ... Steve Lord emphasized the difficulty of the
task because many of these observations are not obtained at the synoptic
times..." [@elsberry1992there]

> "In the modern era, forecasters rely on purely dynamical models that integrate
the Navier-Stokes equations for atmospheric motions to represent both the storm
itself and its surroundings. Numerical models are structured on a computational
grid with temperature, moisture, and wind tabulated in more-or-less rectangular
cells." [@willoughby2007hurricane]

> "Predicting the weather required a widespread but highly coordinated team of
observers and forecasters. Organization was key... They transmitted data from
the field using forms and protocols adapted from those that Abbe had developed
while forecasting the weather from Cincinnati, Ohio. The information arrived by
telegraph coded to maximize accuracy and minimize the word count and changes.
The reports piled up in a rush at the appointed hours." [@willis2006cleveland]

> "Three times daily the duty forecaster broadcast a synopsis of prevailing
weather conditions nationwide..." [@willis2006cleveland]

> "The requirement of synoptic data collection is easily met by the
'conventional' observing system now serving as the main source of meteorological 
data throughout the world, since it is based on human observations which could 
be scheduled at convenient synoptic times." [@morel1971initialization]

Since large-scale weather features are very larger, relevant data for a synoptic
map often comes from locations with different local times. The need for
synchronized weather data collection therefore actually played an important role
in establishing standardized time zones in the United States in [year]. Synoptic
weather times are typically set based on the Universal Time Constant, with times
recorded in Coordinated Universal Time (i.e., Greenwich ...), and time stamps on
weather maps and data based on this time zone often is given a "Z" (for "Zulu",
an earlier name for this timezone) after the time.

> "Accuracy in forecasting depended on the accurate timing of observations. At
first weather observers used local time because no system of standard time
existed, either in the United States or abroad. Local time often meant railroad
time, and that could vary as much as 20 minutes from the time proper to the
local meridian. Abbe called for change, emphasizing the need for simultaneous
readings by Signal Service observers in order to construct accurate charts and
weather maps. In 1875, he recommended to the president of the American
Meteorological Society the establishment of a committee on standard time. ...
Abbe's report in 1879 recommended the system of meridians that is now in general
use ... a system of time meridians one hour apart across the country, anchored
on the Greenwich prime meridian." [@willis2006cleveland]

> "From Abbe's perspective, however, the full potential for dynamic weather
prediction would not be achieved without standard time worldwide."
[@willis2006cleveland]

> "The time in all hydrometeorological products is expressed according to 
a single standard, which is the Universal Coordinated Time (UTC) 
(formerly known as Greenwich Mean Time [GMT] or Zulu (zero) Time [Z])."
[@usnwstime]

> "The U.S. National Weather Service base all data and product times on the 
UTC standard. ... A common practice is to include a time of the observation 
or event denoting the time with a following capital letter 'Z' to indicate the 
time as UTC." [@usnwstime]

> "The extent and the hour of the observations changed over time. Originally the
observers took the readings at 0735, 1635, and 2335, Washington time, but the
last reading was soon changed to 2300 so that the information could be included
in the morning newspapers. Before the introduction of standard time and time
zones within the United States, a confusing multiplicity of local times existed.
In 1870 there were over 100 such regional times. For the purposes of the
meteorological observations, observers used Washington time until 1885.
Thereafter the observers took their readings according to eastern time, or that
at the 75th meridian west of Greenwich, England. Observers also made local-time
readings from 1876 to 1881.[30]" [@raines1996getting]

> "Once the observers had gathered the weather data, the means of reporting and
disseminating it became most important. Like the Smithsonian, the chief signal
officer made arrangements with the leading commercial telegraph companies to
carry the tri-daily reports. Civilian telegraph experts established special
circuits routed to the Signal Office. The initial arrangement with Western Union
regarding transmission was only temporary, and at the end of the trial period
the company refused to continue service. The House Appropriations Committee held
hearings over the dispute and ruled that the company had a mandate to transmit
the weather information as government business. The Signal Corps compensated the
company, however, at rates determined by the postmaster general.[34]" [@raines1996getting]

> "When making the daily telegraphic reports, the weather observers used special
codes to reduce their length to twenty words in the morning and ten in each of
the other two reports, thereby saving the government both time and money.[35]
Regular transmission of the reports began at 0735 on 1 November 1870 from
twenty-four stations stretching from Boston, Massachusetts, south to Key West,
Florida, and west to Cheyenne in the Wyoming Territory. In addition to the
station atop Mount Washington (opened in December 1870), the Signal Corps soon
reached new heights in weather reporting with the station on Pikes Peak that
began reporting in November 1873.[36]" [@raines1996getting]

> "To provide a picture of weather conditions across the country, the observers
made their reports as nearly simultaneous as possible. The weather service did
not initially make forecasts, and the enabling legislation did not specifically
call for it to do so. Eventually general forecasts, referred to as
probabilities, emanated from the Signal Office in Washington. Locally, the
observers posted bulletins and maps in the offices of boards of trade and
chambers of commerce to provide weather information to the public. Post offices
also displayed daily bulletins, and observers supplied local newspapers with
data. Some communities appointed meteorological committees to confer with the
chief signal officer and to serve as a check upon the operations of the local
weather station. On the national level, the Signal Office in Washington issued
daily weather maps compiled from the reports received from all the stations. It
also published the Daily Weather Bulletin, Weekly Weather Chronicle, and the
Monthly Weather Review. All were available for sale to the public. Myer
estimated that through these various means at least one third of American
households received the Signal Corps' weather information in some form. A
railway bulletin service, initiated in 1879, enabled stations along many major
railroads to display weather information.[37]" [@raines1996getting]

> "The Signal Corps' telegraph network soon expanded beyond the eastern
seaboard. In 1874 Congress enacted legislation directing the War Department to
build lines to connect military posts and protect frontier settlements in Texas
against Indians and Mexicans. Other acts authorized lines in Arizona and New
Mexico and, somewhat later, the Northwest. These lines were intended to serve
sparsely settled areas where commercial lines were not yet available. For the
most part, soldiers maintained and operated the lines, but the Corps employed
some civilians. As the telegraph extended its reach, the weather system also
grew, because the operators doubled as weather observers." [@raines1996getting]

> "Until 1934 the Weather Bureau offices operated 12--15 hours a day with two
basic observations taken at 8 a.m. and 8 p.m. The observations were transmitted
via telegraph. There were no satellite images and few upper air observations."
\url{https://www.weather.gov/dvn/armistice_day_blizzard}
